---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# spanish2

<!-- badges: start -->
<!-- badges: end -->

The goal of spanish2 is to allow users to convert from numbers (or their string representations) to their Spanish words equivalents. That is, convert from `100` to "cien", and from `-1325` to "menos mil trecientos veinte y cinco".

I developed it because I frequently needed to convert quantities (usually money) to their text representations in Spanish. There's already [a package for that](https://github.com/rOpenSpain/spanish), but its `to_words()` function has some limitations and produces some unexpected results (the largest number it can handle is 999,999,999). That package has some other interesting features, that `spanish2` lacks, such as converting Spanish text to numbers, so I encourage the user to check that package, too.

**`spanish2` uses long scale**, which is the most frequently used in the Spanish language. That means that 1e9 is **not** "un billón": it is "mil millones". "Un billón" will be 1e12!

## Installation

You can install the development version of spanish2 like so:

``` r
devtools::install_github("pavodive/spanish2")
```

## Examples

The main (and only function) of `spanish2` is `to_words()`, used to convert from numbers to Spanish words:

```{r example}
library(spanish2)

to_words(31262668)
```

It an handle very big **integers**:

```{r example2}
big_number = round(pi * 1e21, 0)
to_words(big_number)
```

But numbers above 1e22 trigger a warning because of wrong arithmetic of big numbers in R: converted text may contain errors:

```{r example3}
bigger_number = round(pi * 1e22, 0)
to_words(bigger_number)
```

Very large (up to 60 digits) can be provided to the function, as long as they are provided as a string:

```{r example4}
set.seed(1)
biggest_number = sample(0:9, 60, TRUE) |>
  paste(collapse = "")

biggest_number

to_words(biggest_number)
```

`spahish2::to_words()` can take a negative number, either as an integer or as a string:

```{r example5}
to_words(-1234)

identical(to_words(-1234), to_words("-1234"))
```

## Simplicity of the Code vs. Style in the Absence of a Standard

There is not (or at least I did not find) a standard transliteration of numbers in Spanish. It seems to vary from region to region. While some countries seem to prefer some forms, others prefer different ones (like 1100: "mil cien", or "mil ciento").

I took side with the simplicity of the code, which may have the consequence of a style that may not be the one favored in your country / region or that of your intended readers or users. The Spanish language is complex and alive, so I hope the simplicity vs. style decision I made does not cause you any significant problems.

What I mean by this "simplicity vs. Style" decision? See for yourself:

| Number | Style 1         | Other Styles              | spanish2's output |
|:------:|:---------------:|:-------------------------:|-------------------|
| 77     | Setenta y siete |                           | setenta y siete   |
| 16     | Dieciséis       | Diez y seis               | diez y seis       |
| 27     | Veintisiete     | Veinte y siete             | veinte y siete    |
| 1100   | Mil cien        | Mil ciento / mil y ciento | mil cien                  |

In the same vein, if you read the code you will see that there is a long function that "cleans the text" (makes some corrections like converting the string "cincodiez" to "cincuenta"). **I am well aware that there is some black magic regex that can reduce the function from its actual 83 lines to some 10 or 15**. But I preferred the readability over the size, so please bear with me (specially if you plan your contribution around this very issue).

## Errors, Bugs and Improvements

Even though I checked it carefully against a good array of numbers, there may be some bugs or errors in the code. For sure there will be ways to improve it or its documentation. Please rise an issue or write a pull request, I'll be happy to review it.
